---
title: "Project: Hockey Profile of a Rude Player" 
author: Giada Brose, Jacob Kasten, Jack Jay
format:  
  html: 
    embed-resources: true 
    code-tools: true 
    toc: true 
    html-table-processing: none 
    theme: slate 
editor: source 
execute:  
  error: true 
  echo: true 
  message: false 
  warning: false
---

# Lab Introduction

The National Hockey League (NHL) is a North American (Canada & United States) professional sporting league with 32 teams. Teams consist of 23 rostered players (20 playable, 3 reserved), with 6 players (1 Goalie, 2 Defensemen, 3 Offensemen) allowed on the ice at any given moment.

Lots of data is recorded by the NHL, including 1. player demographic data, 2. play-by-play data during games, and 3. penalty data. This lab seeks to analyze NHL data through visualization, interpretation, and simulation. In particular, this lab will be structured as though **YOU** are an NHL Data Analyst, where you will find the player who is penalized the most, then use simulation, shot maps, and data cleaning to create a real NHL profile of this **rude** player!

::: callout-tip
Data dictionaries are crucial to understanding how the NHL records. Please use the following data dictionary to aid in your NHL Data Analyst journey: \[https://cpslo-my.sharepoint.com/:w:/g/personal/jkasten_calpoly_edu/IQAPWSaxak0CSaYkmgiRKu4xAYQB_1qvRaAPBzzcpU9tYhw?e=B8PagT \]
:::

# Pre-Work

#### Loading packages

There are several packages needed in order to use this data. Load them here! Additionally, you will need to install some github packages (hockeyR, sportyR) for this lab. Install them using information from \[here\] and \[here\]

```{r}
library(devtools)
library(hockeyR)
library(tidyverse)
library(sportyR)
library(ggimage)
library(knitr)
library(kableExtra)
library(broom)
```

#### Reading in the data

Additionally, we need to read in our .csv hockey datasets!

```{r}
nhl_data <- read_csv("NHL Player and General Data.csv")

pbp_data <- read_csv("PBPData.csv")
```

# Cleaning

The multi-billion dollar NHL cannot afford to have clean data, so we need to clean it for them!

#### Question 0: Cleaning play-by-play data

::: callout-important
The play-by-play dataset is huge, so we've cleaned it for you! However, the second dataset of this lab (the NHL Player Data) will be YOUR responsibility to clean :)
:::

### Question 1: To help simplify and clean the dataset for analysis, complete the following tasks by making a new dataframe, let's call it "nhl_clean":

#### 1A. Keep only the variables that will be used in the analysis: namefirst, namelast, penaltyminutescareer, penaltyminutes2019season, pimyr, team, teamcode,

#### 1B. Create a “fullname” column combining “namefirst” and “namelast” with a space in-between the names, in addition to that, make sure that the names are properly capitalized

#### 1C. Drop “namefirst” and “namelast” after creating "fullname"

```{r}
# Code to do questions 1A-1C
nhl_clean <- nhl_data |>
  select(namefirst, namelast, penaltyminutescareer, 
         penaltyminutes2019season, pimyr, team, teamcode, age, height, weight, birthplace, position, captaincode) |>
  mutate(fullname = str_c(namefirst, namelast, sep = " "),
          fullname = str_to_title(fullname)) |>
  select(-namefirst, -namelast)
```

# SIMULATION

In this section, you will be simulating some expected inferential statistics about the variable 'penaltyminutes2019season' ::: {.callout-tip} There are lots of ways penalty minutes are dished out in hockey! Typically, we see some variation of 3 core types: - Minors, where players receive a 2 minute penalty and their team is short a player on the ice - Major, where players receive a 5 minute penalty and their team is short a player on the ice - Misconducts, where players receive a 10 minute penalty and their team is **NOT** short a player on the ice

This might help contextualize some of this data! :::

### Question 2: Find the minimum, maximum, mean, and standard deviation penaltyminutes2019season. Format it nicely with kable()

```{r}
nhl_clean |> 
  select(penaltyminutes2019season) |> 
  summarise(max_pim = max(penaltyminutes2019season, na.rm = T),
         min_pim = min(penaltyminutes2019season, na.rm = T),
         mean_pim = mean(penaltyminutes2019season,
                         na.rm = T),
         std_pim = sd(penaltyminutes2019season, na.rm = T)) |> 
  kable()

  
```

### Question #3: Simulate a normal distribution using the penaltyminutes2019season variable. Then, find some other interesting statistics to create a box-and-whisker plot for

#### 3A. Simulation, using at least n = 1000.

```{r}
sim_pim <- rnorm(n = 1000, mean = 4.507, sd = 5.312)
head(sim_pim, n=5)
```

#### 3B. Create a histogram of your simulated values

::: callout-tip
You are going to have to turn `sim_pim` into a dataframe to use ggplot. Think about how you can use data.frame() to do this!
:::

```{r}
# there is a dataframe issue here. Same with question 3D
sim_df <- data.frame(sim_pim = sim_pim)

ggplot(sim_df, aes(x = sim_pim)) +
  geom_histogram(bins = 4, color = "white")
```

#### 3C. Find the 25th, 50th, and 75th percentiles for penaltyminutes2019.

```{r}
twentyfifth <- qnorm(0.25, mean = 4.507, sd = 5.312)
fifty <- qnorm(0.5, mean = 4.507, sd = 5.312)
seventyfifth <- qnorm(0.75, mean = 4.507, sd = 5.312)
```

#### 3D. Create a box-and-whisker plot for penaltyminutes2019 using ggplot2 graphics.

```{r}
ggplot(nhl_data$penaltyminutes2019, aes(x = sim_pim)) + 
  geom_boxplot()
```

### Question 4: Find what player has the highest penalty minutes in the 2018-2019 season. Additionally, find their team, age, height, weight, birthplace, position, and captain code, and put it in a nicely formatted table with kable(). Call your new created vector 'rude_player'

```{r}
rude_player <- nhl_clean |> 
  slice_max(penaltyminutes2019season) |> 
  select(fullname, team, age, height, weight, birthplace, position, captaincode) |>
  kable()


kable(
  rude_player,
  caption = "Player with the Highest Penalty Minutes in 2018-2019"
)
```

# Data Analysis

Here, you will be using this data to conduct some classic statistical tests to get a better understanding of hockey penalties. **Maybe this will help you understand hockey penalties more!**



### Question 5a: Create a linear regression model based on the `penaltyminutes2019season` and `weight` variables as seen in question #6.

::: callout-hint
Use the tidy() function to make a clean and easy to read data frame output of our model information!
:::

```{r}
weightlm <- lm(penaltyminutes2019season ~ weight,
             data = nhl_clean)

# Broom output
tidy(weightlm)
```

###Question 5b: Now, create a graph of this relationship using ggplot & broom packages.

```{r}
ggplot(nhl_clean, aes(x = weight, y = penaltyminutes2019season)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

# Shot Map

Shot maps are a unique way to visualize how real NHL games are played! The following questions will guide you on making your very own shot maps. ::: {callout-warning} Make sure you have **ALL** applicable packages loaded :::

### Question 6: Find which game our rude player has the most play-action in the play-by-play data.

```{r}
pbp_data <- pbp_data |>
  mutate(game_date = as.Date(date_time))

rude_player_game <- pbp_data |>
  count(game_id, game_date, home_name, away_name, sort = TRUE) |>
  slice(1)

rude_player_game |>
  kable(
    caption = "Game where Cody McLeod had the most play-by-play action",
    col.names = c("Game ID", "Date", "Home Team", "Away Team", "Number of Events")
  )
```

### Question 7: Now that you have found which game our rude player had the most action in, create a shot map that shows where the players shot at the goal from the ice.

::: callout-hint
Practice using the documentation from the following website:

https://hockeyr.netlify.app/

You don't need to change much about the code, pay close attention to the first few lines of code 🙊
:::

```{r}

# get single game
game <- pbp_data |> 
  filter(game_date == "2018-11-13" & home_abbreviation == "NYR")

# grab team logos & colors
team_logos <- hockeyR::team_logos_colors %>%
  filter(team_abbr == unique(game$home_abbreviation) | team_abbr == unique(game$away_abbreviation)) %>%
  # add in dummy variables to put logos on the ice
  mutate(x = ifelse(full_team_name == unique(game$home_name), 50, -50),
         y = 0)

# add transparency to logo
transparent <- function(img) {
  magick::image_fx(img, expression = "0.3*a", channel = "alpha")
}

# get only shot events
fenwick_events <- c("MISSED_SHOT","SHOT","GOAL")

shots <- game %>% filter(event_type %in% fenwick_events) %>%
  # adding team colors
  left_join(team_logos, by = c("event_team_abbr" = "team_abbr"))

# create shot plot
geom_hockey("nhl") +
  ggimage::geom_image(
    data = team_logos,
    aes(x = x, y = y, image = team_logo_espn),
    image_fun = transparent, size = 0.22, asp = 2.35
    ) +
  geom_point(
    data = shots,
    aes(x_fixed, y_fixed),
    size = 6,
    color = shots$team_color1,
    shape = ifelse(shots$event_type == "GOAL", 19, 1)
    ) +
  labs(
    title = glue::glue("{unique(game$away_name)} @ {unique(game$home_name)}"),
    subtitle = glue::glue(
    "{unique(game$game_date)}\n
    {unique(shots$away_abbreviation)} {unique(shots$away_final)} - {unique(shots$home_final)} {unique(shots$home_abbreviation)}"
    ),
    caption = "data from hockeyR | plot made with sportyR"
    ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    plot.caption = element_text(hjust = .9)
    )
```

### Question 8: Now that you know how shot map creation works, make a function that will create a shot map for any game we would like. Make the function take a `date` argument, along with a `hometeam` argument. Name your function `getshotmap`. Test your function with with the 2018-10-03 TOR prompt below this question!

```{r}
getshotmap <- function(date, hometeam){

# get single game
game <- pbp_data |> 
  filter(game_date == date & home_abbreviation == hometeam)

# grab team logos & colors
team_logos <- hockeyR::team_logos_colors %>%
  filter(team_abbr == unique(game$home_abbreviation) | team_abbr == unique(game$away_abbreviation)) %>%
  # add in dummy variables to put logos on the ice
  mutate(x = ifelse(full_team_name == unique(game$home_name), 50, -50),
         y = 0)

# add transparency to logo
transparent <- function(img) {
  magick::image_fx(img, expression = "0.3*a", channel = "alpha")
}

# get only shot events
fenwick_events <- c("MISSED_SHOT","SHOT","GOAL")

shots <- game %>% filter(event_type %in% fenwick_events) %>%
  # adding team colors
  left_join(team_logos, by = c("event_team_abbr" = "team_abbr"))

# create shot plot
geom_hockey("nhl") +
  ggimage::geom_image(
    data = team_logos,
    aes(x = x, y = y, image = team_logo_espn),
    image_fun = transparent, size = 0.22, asp = 2.35
    ) +
  geom_point(
    data = shots,
    aes(x_fixed, y_fixed),
    size = 6,
    color = shots$team_color1,
    shape = ifelse(shots$event_type == "GOAL", 19, 1)
    ) +
  labs(
    title = glue::glue("{unique(game$away_name)} @ {unique(game$home_name)}"),
    subtitle = glue::glue(
    "{unique(game$game_date)}\n
    {unique(shots$away_abbreviation)} {unique(shots$away_final)} - {unique(shots$home_final)} {unique(shots$home_abbreviation)}"
    ),
    caption = "data from hockeyR | plot made with sportyR"
    ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    plot.caption = element_text(hjust = .9)
    )

}

```

#### Question 8 Test

```{r}
getshotmap("2018-10-03", "TOR")
```

# Player Profile

We have so much data about our beloved rude player! Its now time to create a professional player profile!

### Question 9: Create a neatly organized player profile for our rude player! Include:

-   a nicely formatted table with their name, age, height, birthplace, position, and team.
-   a photo of this player
-   a shot map of this player's most active game
-   an interesting statistic about high penalty minutes

```{r}
# Rude player profile
rude_player_profile <- nhl_clean |>
  slice_max(penaltyminutes2019season, n = 1, with_ties = FALSE) |>
  select(fullname, age, height, birthplace, position, team, penaltyminutes2019season)

rude_player_profile |>
  kable(
    caption = "Player Profile: NHL's Rudest Player",
    col.names = c("Name", "Age", "Height", "Birthplace", "Position", "Team", "Penalty Minutes")
  ) |>
  kable_styling(full_width = FALSE)


knitr::include_graphics(
  "https://cms.nhl.bamgrid.com/images/headshots/current/168x168/8469457.jpg"
)


pim_cutoff <- quantile(nhl_clean$penaltyminutes2019season, 0.95, na.rm = TRUE)

nhl_clean |>
  summarise(
    rude_player_pim = rude_player_profile$penaltyminutes2019season,
    league_average_pim = mean(penaltyminutes2019season, na.rm = TRUE),
    top_5_percent_cutoff = pim_cutoff,
    rude_player_is_top_5_percent = rude_player_pim >= top_5_percent_cutoff
  ) |>
  kable(
    digits = 2,
    caption = "How unusual are this player's penalty minutes?"
  ) |>
  kable_styling(full_width = FALSE)
```
